Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-07-04 paper/main.tex
all the spivak fiber goodness

\subsection{Equivariant Transformations}
Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the column types in a table, when they translate data to visual representations. The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure. Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}. This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}. The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}. They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

@@

A fiber bundle $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation}

 $\textcolor{set}{\Gamma}$ denotes the set of all sections. The set of sections $\dsection$ of a bundle are continuous functions 
\begin{equation}
  \label{eq:atct:fb:sections}
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\} 
\end{equation}

\subsubsection{Topology: Base Space \dbase}
We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it consists of the set of points ${\dbasepoint \in dbasec}$ and topology $\mathcal{T}_{\dbase}$. A topology $\mathcal{T}_{\dbase}$ is a collection of open sets $\openset$ such that the empty set and \dbase are in $\mathcal{T}_{\dbase}$,and every union and intersection of opensets in $\mathcal{T}_{\dbase}$ are also in $\mathcal{T}_{\dbase}$ \cite{bradleyTopologyCategoricalApproach2020}. An openset \openset\ is a generalization of an open interval to a region, meaning an open set contains points inside a bounded region on \dbase and excludes the points on the boundary \cite{TopologyVsTopology}. A property of the base space \dbase\ is that it is a quotient topology, ,meaning it has the finest topology of $\dbase$, which  means that it contains the largest possible collection of distinct open sets on $\dbase$ \cite{QuotientSpaceTopology2020}. For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}. The topology on $\dbase$ encodes the space of smallest possible subsets of the data, and the union and intersection rules express that these subsets must be combinable into larger subsets of the data.

We encapsulate the notion of a programmatic topology type as the base space category $\mathcal{\dbasec}$. The base space category $\mathcal{\dbasec}$ has open set objects $\opensetc$. The morphisms of the category are inclusions $\iota: \opensetc_1 \rightarrow \opensetc_2$ such that $\opensetc_1 \subseteq \opensetc_2$. By definition of a catagory\cite{barrCategoryTheoryComputing}, morphisms are commutative $(\iota_{1,2}: \openset_{1} \rightarrow \openset_{2}) \circ (\iota_{2,3}: \openset_{2} \rightarrow \openset_{3}) = \iota_{1,3}: \openset_{1} \rightarrow \openset_{3}$. Another property of morphisms is that they are associative such that $\iota_{3,4}\circ(\iota_{2,3}\circ\iota_{1,2}) = (\iota_{3,4} \circ \iota{2,3})\circ \iota_{1,2}$ where $\iota_{3,4}:\openset_{3}\rightarrow \openset_{4}$. Each object also has an identity morphism $1_{\openset_i}: \openset_i \rightarrow \openset_i$. We propose that these categorical types can be mapped to 
programmatic types that describe the data topology and that the inclusion morphisms can be used to assert that the continuity has been preserved when defining subsets of data.

\subsection{Field Types: Fiber Space \dfiber}
We encode the field types of a dataset as the  \textcolor{fiber}{fiber space} \dfiberc of the fiber bundle because it allows us to encode fields of varying types and dimensionalities in a consistent manner. The fiber \dfiber\ is the preimage of the projection function $\pi$ at a point in the base space $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$, meaning that all points in a fiber space $\dfiber_{\dbasepoint}$ map to the same point $\dbasepoint$. The fiber space can be used to encode the space of all possible values of a data field and the fibers can be named, as described by Spivak \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}. The space of all possible values is analogous to a CS data type; therefore the fiber space can be used to encode the data types of the data fields. The fiber is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields.

As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data, which is akin to a schema in a database \cite{ullmanFirstCourseDatabase2008}. The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type. The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$. The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$. The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values. For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$. The $\otimes$ encapsulates both the sets associated with each fiber $\mathbb{R} \times \mathbb{R} = \mathbb{R}^{2}$ and the morphisms associated with each functor $(\dfunct_{speed}, \dfunct_{direction}) = \dfunct_{wind}$. Defining the field schema as a monoidal category allows us to formally express the structure of the field, as defined by its sets and functions, at different levels of composition as needed by different visualization tasks.

\subsection{Data: Section \dsection}
We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition. Sections  $\dsection: \dbasepoint \mapsto \delement$ map from a point in the basespace $\dbasepoint \in \dbase$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$. We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$. Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$ which encapsulates the notion of multiple datasets with a shared fiber and base space. For example, all functions on the interval [0, 1] or a set of database tables with the same schema.

@@

\subsection{intrepretation/example}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{fbnet.png}
  \caption{\label{fig:atct:networkbundle}}
\end{figure}
\note{figure out picture/example}
Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records. Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types. For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$. The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}. She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset. She defines the value as a dependent attribute. In this work we use a more general notion of the topological space as an indexing space for the data records. For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized. Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.

\subsection{Uniform Abstract Graphic Representation}
Fiber bundles are such a generalizable data abstraction that we can also use them to encode the output of a visual bundle; bundles allow for fully expressing the properties of the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) data to graphic transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.

@@

Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
A presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
The presheaf expresses that every subset of the data $\openset$ has an associated set of data sections $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$. The presheaf is a contravariant functor because the morphisms between input objects
\begin{equation*}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
go in the opposite direction from the restriction morphisms $\iota^*$ between the sets
\begin{equation*}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. We use a  presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$.

The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space. The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$. The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data. In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.

@@

\subsection{Mathematical Structure of Data}
\label{sec:atct:sheaves:measurments}
Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. We generalize binary operations to actions $\dfunc \in \dfuncset$ on elements of the bundle $\delement \in \dtotal$. This is expressed with a composition function  $act: \dfuncset \times E \rightarrow \dtotal$ that applies the action to an element of \dtotal\ such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$. We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.

>>>>

\subsection{Data Field Structure}
\label{sec:related-work:equivariance}
Along with preserving topology, visualization algorithms are expected to preserve the structure of the data fields, i.e. the column types in a table, when they translate data to visual representations. The mathematical term for this is equivariance and an equivariant map is a map that transforms input with structure to output with compatible structure. Bertin codified the idea that visual encoding of a data field should match the structure of the field; for example ordered data should be encoded as an ordered visual element or quantitative data as a quantitative visual element \cite{bertinSemiologyGraphicsDiagrams2011a}. This structure mapping is necessary to ensure that the visual representations in graphs are not distorting the data\cite{tufteVisualDisplayQuantitative2001} and graphs are easier to understand when properties match \cite{norman_things_smart}. The expectation of matching structure is generalized as a mapping of a binary operator that acts on the data domain to a binary operator that acts on the graphic domain by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. An invertible binary operator is the basis of Kindlemannn and Scheideggers's algebraic model of visualization\cite{kindlmannAlgebraicProcessVisualization2014}. They propose three specific types of equivariance: that the visualization should not change if the data representation (i.e. the data container) changes, that data maps unambiguously to visual elements \cite{ziemkiewiczEmbeddingInformationVisualization2009}, that changing the data should correspond to changes in the visualization in a perceptually significant manner. In \autoref{sec:atct:sheaves} we introduce a uniform abstract data representation layer, the expectation of unambiguous mappings is expressed in \autoref{sec:atct:xi}, and the definition of equivariance introduced in \autoref{sec:artist:equivariance} includes the correspondence between data and visual changes. In this work, we formally specify the conditions necessary to build structure preserving components using category theory, because it provides a rich language for expressing the structure of objects, functions between objects, and the constraints that must hold for these functions to compose \cite{wielsManagementEvolvingSpecifications1998,yorgeyMonoidsThemeVariations}. A brief visualization oriented introduction to category theory is in Vickers et al \cite{vickersUnderstandingVisualizationFormal2013}, but they are applying category theory to semantic concerns about visualization design rather than library architecture.

@@

\begin{definition} A \textbf{fiber bundle} $(\dtotalc, \dbasec, \pi, \dfiberc)$ is a structure with topological spaces $\dtotalc, \dfiberc, \dbasec$ and continuous surjective map $\pi: \dtotalc \rightarrow \dbasec$ \cite{FiberBundle2020,spanier1989algebraic}. 
\begin{equation}
  \label{eq:atct:fb:intro}
  \begin{tikzcd}[ampersand replacement=\&, row sep=huge]
   \dfiberc
    \arrow[r, hook, color=total] \& 
    \dtotalc
    \arrow[d, "\pi"',color=total] \\
     \& 
  \dbasec
     \arrow[u, "\dsectionc"', bend right, pos=.5, color=section, dashed]
  \end{tikzcd}
\end{equation} 
A map $\pi$ is a bundle projection map when 
\begin{enumerate}
  \item the \textcolor{fiber}{fiber space} \dfiberc\ is the preimage of the projection function $\pi$ at a point $\dbasepoint$ in the \textcolor{base}{base space} \dbasec\ such that $\dfiber_{\dbasepoint} = \pi^{-1}(\dbasepoint)$
  \item there is an open neighborhood $\openset_{\dbasepoint}$ surrounding each point in the base space $\dbasepoint \in \openset_{\dbasepoint} \subset \dbase$ such that the \textcolor{total}{total space} \dtotalc\ over the neighborhood is locally trivial $\dtotal\restriction_{\openset} = \openset \times \dfiber$
\end{enumerate}
\end{definition}
The first condition for a fiber bundle means that all points in a fiber space over a point in the base space map to the same base space point, which we rely on to establish that there is a formal binding between typed data value, discussed in \autoref{sec:atct:fiber:fiber} and the corresponding index, discussed in \autoref{sec:atct:fiber:base}. The condition of local triviality means that the fibers are identical over open neighborhoods and is what allows fiber bundles to be extremely generalizable, as discussed in \autoref{sec:atct:fiber:sections}

\begin{definition} A \textcolor{section}{\textbf{section}} $\dsectionc: \dbasec \rightarrow \dtotalc$ over a fiber bundle is a smooth right inverse of $\pi$ such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$ for all $\dbasepoint \in \dbase$
\end{definition}
In \autoref{sec:atct:fiber:sections}, we propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner.

\subsubsection{Topological Structure: Base Space \dbase}
\autoref{sec:atct:fiber:base}
We encode the underlying connectivity of the data as the \textcolor{base}{base space} \dbase\ of the fiber bundle because it provides a uniform language for describing the topology regardless of how the data is subset. The base space \dbase\ is a topological space $(\dbase,\mathcal{T}_{\dbase})$, which means it  consists of the set of points ${\dbasepoint \in \dbase}$ and topology $\mathcal{T}_{\dbase}$ \cite{munkresElementsAlgebraicTopology1984}. The set \dbase\ is equipped with open subsets \openset, which are a generalization of an open interval. This means opensets contain points inside a bounded region on the topological space and excludes the points on the boundary \cite{TopologyVsTopology}. An openset that surrounds a point \dbasepoint\ is called an open neighborhood $\openset_{\dbasepoint}$. Opensets are key to constructing topologies, and are used in this paper to model subsets of the data contextualized within the larger continuity those subsets belong to.

\begin{definition}
A \textbf{topology} $\mathcal{T}_{\dbase}$ on a space \dbase\ is a collection of open sets \openset\ of \dbase\ that satisfy the properties 
  \begin{itemize}
    \item the empty set $\varnothing$ and $\dbase$ are in $\mathcal{T}_{\dbase}$
    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the intersection $\openset_i \cap \openset_j$  is in $ \mathcal{T}_{\dbase}$
    \item for all $\openset_i, \openset_j \in \mathcal{T}_{\dbase}$, the union $\openset_i \cup \openset_j$ is in $ \mathcal{T}_{\dbase}$
  \end{itemize}
  such that the the topology provides an axiomatic way to construct the topological space $\dbase \coloneqq (\dbase,\mathcal{T}_{\dbase})$\cite{bradleyTopologyCategoricalApproach2020}.
\end{definition}
 A property of the base space \dbase\ of a fiber bundle is that it is a quotient topology, which  means that it contains the largest possible collection of distinct opensets, meaning each openset is as small as possible, of $\dbase$ that still preserve the underlying continuity of \dbase \cite{QuotientSpaceTopology2020}. For our purposes, $\dbase$ is a set of points that encodes the underlying connectivity of the our dataset, for example, $\dbase$ could be the discrete points, interval, or plane in \autoref{fig:related-work:continuity:ktypes}. We express the allowable ways to subset and compose points such that the connectivity is preserved as the topology $\mathcal{T}_{\dbase}$ on the data continuity.

The topological structure of the data can be expressed programmatically by constructing data types that encapsulate the class of topological structure-.i.e point, line, plane, network, etc. We propose that these data types can be formally specified as objects of a category. 
\begin{definition} 
   An arbitrary \textbf{category} $\mathcal{C}$ consists of the following \textit{data}: 
\begin{enumerate}
  \item \textit{object}: X
  \item \textit{identity morphism}: $X \xrightarrow {id_x} X$
  \item \textit{morphisms}: $X \xrightarrow{f} Y$ between objects X, Y in $\mathcal{C}$
  \item \textit{composite morphism}: given the morphisms $X \xrightarrow{f} Y$ and $Y \xrightarrow{g} Z$, there exists a morphism $X \xrightarrow{g\circ f} Z$
\end{enumerate}
The data must satisfy the following \textit{rules}
\begin{enumerate}
  \item \textit{unitality:} for every morphism $ X \xrightarrow{f} Y$, there exists identity morphisms $id_x, id_y$ such that $f \circ id_x = f = id_y \circ f$
  \item \textit{associativity:} if any three morphisms $f, g, h$ are composable, 
    \begin{equation*}
      \begin{tikzcd}
        X \arrow[r, "f"] \arrow[rrr, "h\circ(g\circ f) = (h\circ g)\circ f"', bend right, dashed] & Y  \arrow[r, "g"] & Z \arrow[r, "h"] & W
        \end{tikzcd}}
  \end{equation*}
  then they are associative such that $h\circ(g\circ f) = (h \circ g) \circ f$  \cite{lawvere2009conceptual,riehlCategoryTheoryContext,maclaneCategoriesWorkingMathematician2013,fongInvitationAppliedCategory2019}. 
  \end{enumerate}
\end{definition}

We construct a base space category $\mathcal{\dbasec}$ to encapsulate the topological structure of the data. The category $\mathcal{\dbase}$ has open set objects $\opensetc$ and inclusion morphisms $\opensetc_i \xrightarrow{\iota} \opensetc_j$ such that $\opensetc_i \subseteq \opensetc_j$. The composability property expresses that inclusion is transitive, while associativity expresses that the inclusion functions can be curried in various equivalent groupings. By formally specifying the properties of the topological structure datatypes as $\mathcal{\dbase}$, we can express that these are the properties that are required as part of the implementation of the data type objects.

\subsubsection{Data Fields: Fiber Space \dfiber}
\autoref{sec:atct:fb:fiber}
As mentioned in \autoref{sec:related-work:equivariance}, equivariance is traditionally evaluated as the preservation of field structure, which is based on the field type. To have a rich language for expressing field type, we generalize Spivak's fiber bundle formulation of table \cite{spivakDatabasesAreCategories2010,spivakSIMPLICIALDATABASES}.

\begin{definition} A \textbf{type specification} is the bundle map $\pi: \mathscr{U} \rightarrow \textbf{DT}$. The base space $\textbf{DT}$ is the set of data types, and the total space $\mathscr{U}$ is the set of domains of those data types. Given an arbitrary data type $T \in \textbf{DT}$, the preimage $\pi^{-1} \subset \mathscr{U}$ is the domain of $T$ and $x \in \pi^{-1}(T)$ is an object of type $T$.  
\end{definition}
For example, if $T=\texttt{int}$, then the image $\pi^{-1}(\texttt{int}) = \mathbb{Z} \subset \mathscr{U}$ is the set of all integers. The domain $\mathscr{U}_{{\sigma}}$ is also a topological space, meaning there is no restriction on the dimensionality of the field, thus allowing for the encoding of multi-dimensional fields in the same manner as single dimensional fields. Since many fields can have the same datatype, Spivak formally defines a mapping from field name to field data type, akin to a  data base schema \cite{ullmanFirstCourseDatabase2008}. 
\begin{definition} A \textbf{schema} consists of a pair $(\fnames, \sigma)$ where $\fnames$ is the set of field names and $\sigma: \fnames \rightarrow \textbf{DT}$ is a function from field name to field data type. 
\end{definition}
The function $\sigma$ is composed with $\pi$ such that the bundle  $\pi_{\sigma}:\mathscr{U}_{\sigma} \rightarrow \fnames$ associates a field name $c \in C$ with the corresponding data type domain. Spivak then defines data records as sections on the bundle defined by $\pi_{\sigma}$. 
\begin{definition} A \textbf{record} is a function $\delement: \fnames \rightarrow \mathscr{U}_{\sigma}$ and the set of records on $\pi_{\sigma}$ is denoted $\Gamma^{\pi}(\sigma)$.
\end{definition}
Records must return an object of type $\sigma(\fname) \in \textbf{DT}$ for each field $c \in C$. Spivak then constructs tables as sections $\dsection: \dbase \rightarrow \Gamma^{\pi}(\sigma)$ from a key space $\dspace$ to the set of records $\Gamma^{\pi}(\sigma)$ on the schema bundle. Building on Spivak's construction, we define the fiber space in this work as the set of records 
\begin{equation*}
  \dfiberc \coloneqq \{\delement: \fnames \rightarrow \mathscr{U}_{\sigma} \bigm{\vert} \pi_{\sigma}(\delement(\fname)) = \fname\;for\;all\; \fname \in \fnames \}
\end{equation*}
such that the preimage of a point is the corresponding domain $\pi^{-1}(\dbasepoint) = \dfiber_{k} = \mathscr{U}_{{\sigma}_{\dbasepoint}}$. This construction of the fiber as a schema bundle allows for the unique representation of data fields that have identical domains because they have the same underlying data type while still retaining the richness of a formal typing system.

As with the base space category $\mathcal{\dbase}$, we propose a fiber category $\mathcal{\dfiberc}$ to encapsulate the field types of the data. The fiber category $\mathcal{\dfiberc}$ is a monoidal category, meaning the category has a single object $\dfiberc$ of an arbitrary type. The morphisms on the fiber object are $\dfunctc \in Hom(\dfiberc, \dfiberc)$. The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$. The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values. For example, wind can be represented as two fields $\dfiber_{speed} \times \dfiber_{direction}$ or a composite fiber field $\dfiber_{speed} \otimes \dfiber_{direction} = \dfiber_{wind}$. The $\otimes$ encapsulates both the sets associated with each fiber $\mathbb{R} \times \mathbb{R} = \mathbb{R}^{2}$ and the morphisms associated with each functor $(\dfunct_{speed}, \dfunct_{direction}) = \dfunct_{wind}$. Defining the field schema as a monoidal category allows us to formally express the structure of the field, as defined by its sets and functions, at different levels of composition as needed by different visualization tasks.

\subsubsection{Data: Section \dsection} 
\autoref{sec:atct:fb:sections}
We encode data as a \textcolor{section}{section} \dsectionc\ of a bundle because this allows us to incorporate the topology and field types in the data definition. We can define these section functions locally, meaning over a specific open subset \openset\ of \dbase\ 
\begin{equation}
  \label{eq:atct:fb:sections}
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\opensetc} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;for\, all\; \dbasepointc \in \opensetc \big\} 
\end{equation}
such that each section function $\dsection: \dbasepoint \mapsto \delement$ maps from a point $\dbasepoint \in \openset$ to a record in the fiber space $\delement \in \dfiber_{\dbasepoint}$ over that point. Bundles can have multiple sections, as denoted by $\Gamma(\openset, \dtotal\restriction{\openset})$. We can therefore model data as structures that map from an index like point $\dbasepoint$ to a data record $\delement$, and encapsulate multiple datasets with the same fiber and base space as different sections of the same bundle.

@@

\subsubsection{intrepretation/example}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{fbnet.png}
  \caption{\label{fig:atct:networkbundle}}
\end{figure}
\note{figure out picture/example}
Extending Spivak's fiber bundle description of a table, we define the values returned by a section as named and typed records. Since the fiber space is a topological space, it does not really have restrictions on the dimension of the types it can encode or the structure of those types. For example, wind has a speed component and a direction component and we can consider those separately $\dfiber_{speed} = \mathbb{R}, \dfiber_{direction} = \mathbb{R}$ or as one fiber$\dfiber_{wind} = \mathbb{R}^{2}$. The base space and the section evaluated on the base space form a sort of key-value pair, as described by Munzner \cite{munznerVisualizationAnalysisDesign2014}. She imbues semantic meaning to the topology, wherein the key is an independent attribute used as a look up key to retrieve the value in the dataset. She defines the value as a dependent attribute. In this work we use a more general notion of the topological space as an indexing space for the data records. For example, a data cube of temperature can be encoded such that every fiber is a set of 2D temperature surfaces $\dfiber_{tempsurf}$, the base space is a 1D continuous interval, and each section returns a surface $\dsection(\dbasepoint) = f(x,y)$ where x and y are already parameterized. Alternatively, the fiber can encode each cell of the data cube $\dfiber_{lat} \times \dfiber_{lon} \times \dfiber_{temp}$, the base space could be a 3D volume, and each section returns a $\dsection(\dbasepoint) = \{lat, lon, temp\}$ record.

\subsubsection{Uniform Abstract Graphic Representation}
Fiber bundles are such a generalizable data abstraction that we can also use them to encode the output of a visual bundle; bundles allow for fully expressing the properties of the graphic that must be symmetric to the data in an equivariant (\autoref{sec:related-work:equivariance}) data to graphic transformation. We denote the output as a graphic, but the use of bundles allows us to generalize to output on any display space, such as a screen or 3D print. 
\begin{equation}
  \label{eq:atct:fb:graphic}
  \gfiberc \hookrightarrow \gtotalc \xrightarrow{\pi} \gbasec
\end{equation}
The total space \gtotalc\ is an abstraction of an ideal (infinite resolution) space into which the graphic can be rendered. The base space \gbasec\ is a parameterization of the display area, for example the inked bounding box in cairo \cite{CairographicsOrg}. The fiber space \gfiberc\ is an abstraction of the renderer fields, for example $\{x,\,y,\,r,\,g,\,b,\,a\}$.

@@

\note{add definition of functor + presheaf}
Presheaves are functors, meaning they are functions from objects of one category to objects of another category\cite{WhatFunctorDefinitions} that preserve the morphisms in each category. 
A presheaf is a contravariant functor from an object in an arbitrary category to an object in the category \setb\cite{nlab:presheaf}. 
\begin{equation}
  \label{eq:atct:sheaf:functor}
  \sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
The presheaf expresses that every subset of the data $\openset$ has an associated set of data sections $\cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$. The presheaf is a contravariant functor because the morphisms between input objects
\begin{equation*}
  \iota: \openset_1 \rightarrow \openset_2
\end{equation*}
go in the opposite direction from the restriction morphisms $\iota^*$ between the sets
\begin{equation*}
  \iota^*: \cgamma{\openset_2}{\dtotal_{\openset_2}} \rightarrow \cgamma{\openset_1}{\dtotal_{\openset_1}}
\end{equation*}
which are the outputs of the sheaf $\sheaf(\openset) = \cgamma{\openset}{\dtotal\restriction_{\openset}}$. We use a  presheaf to describe the expectation that any set of sections defined over a larger space $\dsection \in \Gamma(\opensetc_2, \dtotal\restriction_{\openset_{2}})$ is also defined over subspaces of that space $\dsection\restriction_{\opensetc_{1}}$. 
\note{make these formal definitions again}
The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space. The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$. The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data. In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.

@@

\subsection{Mathematical Structure of Data}
\label{sec:atct:sheaves:measurments}
\note{move to Artist }
Equivariance in a visual context was formally defined as the preservation of a binary operator from one domain to another by Mackinlay\cite{mackinlayAutomaticDesignGraphical1987}. We generalize binary operations to actions $\dfunc \in \dfuncset$ on elements of the bundle $\delement \in \dtotal$. This is expressed with a composition function  $act: \dfuncset \times E \rightarrow \dtotal$ that applies the action to an element of \dtotal\ such that $\dfunc(\delement_0) = \delement_1, \delement_0, \delement_1 \in \dfunc$. We generalize to a family of actions because that allows for expanding the set of allowable transformations on the data beyond a single operator.
